---
title: "Managing Connections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Connections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`sqlhelper` can maintain one or many database connections internally. This
article describes how to configure and create connections, and the functions for
inspecting and accessing them.

## Creating connections

You can define database connections in config files which are read and processed
by the `connect()` function. For example, if your config file is called
`my_db_config.yml`, you might run:

``` r
library(sqlhelper)

connect("my_db_config.yml")
```

### Config files

A config file can contain one or more connection definitions.
Connection definitions are [YAML](https://yaml.org/). The following YAML chunk defines a connection called 'dap'. 

```yaml
dap:
  driver_type: sqlserver
  pool: yes
  description: >
    Databases managed by ADD teams on the Data and Analytics Platform
  connection:
    Driver: "{ODBC Driver 17 for SQL Server}"
    Server: "Dap-sql01"
    Trusted_Connection: "yes"
```

The first line is a name for this connection; subsequent lines contain
information about the connection.

The **driver_type** line defines the brand of database, and hence the driver
package that will be used. This parameter is required. Current options are:

* 'odbc' (odbc)
* 'sqlserver' (odbc)
* 'sqlite' (RSQLite)
* 'postgresql' (RPostgres)
* 'mariadb'(RMariaDB)
* 'mysql'(RMariaDB)
* 'bigquery' (bigquery)

The **pool** line determines whether a single connection is required (as
returned by `DBI::dbConnect()`) or a pool of connections (as returned by
`pool::dbPool()`). The latter is recommended for
[shiny](https://shiny.rstudio.com/) applications. This parameter defaults to
'no', meaning a single connection will be provided.

The **description** line provides a short description of this connection. This
parameter can be omitted.

The **connection** section contains several lines which provide the parameters
required for the connection itself, to be passed to the constructor function as
name=value pairs. For example, you will need to supply a `Driver` connection
string if you are using an `odbc` driver. The **server** parameter will probably
always be required but see `DBI::dbConnect()` for more details.

In summary, the elements needed for every connection are:

#. The connection **name**, and
#. The **driver_type**, and
#. A **connection** section containing a **Server** element

The others *may* be needed for some connections or may be omitted entirely. The function `config_examples()` provides several example configurations. 

### The config search path

By default, the `connect()` function reads not only the config file named by its
`config_filename` parameter, but also any config files found on a search path
that includes both the user's and the site's config directories.

If you use the same database connections often, you can place a config file
called `sqlhelper_db_config.yml` in the directory named by
`rappdirs::user_config_dir()`. Similarly, if you are a site admin you can
provide your users with pre-configured connections by placing a config file of
the same name in the directory named by `rappdirs::site_config_dir()`. In either
of these cases, no config filename is needed and `connect()` can be called
without arguments. Nonetheless, in the case where you require additional
connections to those provided by user- and site-wide configurations, you may
provide a `config_filename` argument.

### File precedence, inheritance, and conflicts across config files

Site-wide config files are read first, then user-wide, then named files. If
conflicting connection names or connection elements are encountered between
config files, values from later files overwrite those from earlier. This
operates on an element-by-element basis; for example, if your site-wide config
contains:

```yaml
my_conn:
  server_type: mysql
  connection:
    Server: "organization_server_host"
```

but your named file contains:

```yaml
my_conn:
  description: Private MySQL database
  connection:
    Server: "my_alternate_server_host"
```
then `sqlhelper` will attempt to create a connection described by:

```yaml
my_conn:
  server_type: mysql
  description: Private MySQL database
  connection:
    Server: "my_alternate_server_host"
```

### Suppressing the config search path

If only one config file is wanted, the config search path may be suppressed by
supplying the `exclusive = TRUE` argument.

If `exclusive = TRUE`, then `connect()` will check whether the `config_filename`
is one of 'user' or 'site'. If it is, the user- or site-wide config file will be
sought; if it is not 'user' or 'site', it will be treated as the name of a
config file.

## Browsing the connection cache

After `connect()` has been called, you may inspect the created connections with
`connection_info()`.

```{r}
library(sqlhelper)

conf_fn <- system.file("examples/sqlhelper_db_conf.yml",
                       package = "sqlhelper")

readLines(conf_fn) |>
  writeLines()

connect(conf_fn, exclusive = TRUE)


connection_info()
```


## Accessing individual connections

Connections may be accessed by name, with `live_connection()`

```{r}
myconn <- live_connection("simple_sqlite")
myconn
```

## The default connection

The first connection in the file with the highest precedence (i.e. the last one to be read) becomes the default default. This connection will be used whenever queries or files of SQL are executed (`runqueries()` or `runfiles()`) with without explicitly stating which connection to use. 

You can check the default connection with `connection_info()` or `get_default_conn_name()` and change it with `set_default_conn_name()`.

```{r}
get_default_conn_name()

set_default_conn_name("pool_sqlite")

get_default_conn_name()

get_default_conn_name() |>
  live_connection()
```

The last of these - combining `get_default_conn_name()` with `live_connection()` to obtain the default connection - is common enough to warrant a convenience function, `default_conn()`.

## Checking, closing and re-opening connections

You can check whether connections are live in two ways:

#. `connection_info()` provides a 'live' field for interactive use
#. The functions `is_connected()` and `not_connected()` are intended for programmatic use.

To close all connections and remove them from the internal cache, use `disconnect()`.

Individual connections may be closed with `DBI::dbDisconnect()`. 

`connect()` closes any open connections before reading config files. 


```{r}
conn_name <- "simple_sqlite"

is_connected(conn_name)
connection_info(conn_name)$live

myconn <- live_connection(conn_name)
DBI::dbDisconnect(myconn)

connection_info()

if(not_connected(conn_name)){
  message(glue::glue("{conn_name} is not available, reconnecting..."))
  connect("configs/sqlhelper_db_conf.yml", exclusive = TRUE)
}

connection_info()

disconnect()

connection_info()
```

