---
title: "sqlhelper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sqlhelper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Quickstart

Specify your connections to databases in a yaml file:

```{r}
# load sqlhelper
library(sqlhelper)

# If your admin has kindly provided a site-wide config, or you have set one up 
# in your home directory, you can connect to your all your databases with:
# connect()

# Otherwise, you'll need to provide a config file:
connect("configs/sqlhelper_db_conf.yml")
```

[Managing Connections] (below) and `vignette("connections")` explain how to specify connections.

For data wrangling, you probably want to write SQL queries in their own files and use `runfiles()`:

```R

# write some queries in sql files
files_to_run <- list("path/to/create_temporary_dataset.sql",
                     "path/containing/final_merge.sql",
                     "folder/enclosing/extract_subsets.sql")

# Run the queries and save the results
results <- runfiles(files_to_run)

# Inspect the results. By default, runfiles() returns a list of the results of 
# each query in the files you provided. Results of a specific query can be accessed by the
# the name of the query. See the article 'Executing SQL' for more on named queries.

head(results$my_extracted_subset)
```

For exploratory work you might just want to run some short queries; in this case `runqueries()` is for you:

```R
# write some queries
my_queries <- list(
  usedb="use COVID19", 
  showtabs="select * from INFORMATION_SCHEMA.TABLES" # We're using SQLServer
)

# Run the queries and save the results
results <- runqueries(my_queries)

# Inspect the results. runqueries() returns a list with one element per query.
# You can access them using the names of the queries:
head(results$showtabs)
``````

## Managing Connections

Connections are configured with [YAML](https://yaml.org/). An minimal config file could contain:

```yaml
my_sqlite_conn:
  driver_type: sqlite
  connection:
    Server: ":memory:"
```

See `vignette("connections")` for complete details.

## Running queries

Queries in SQL files can be named using comments, and SQL can be interpolated with R values.

```{r}
library(sqlhelper)

connect("configs/sqlhelper_db_conf.yml")

sqlite_con <- live_connection("simple_sqlite")
DBI::dbWriteTable(sqlite_con,
                  name = "IRIS", 
                  value = iris)

n_longest_petals <- 6

readLines("sql/example.sql") |>
  writeLines()

result <- runfiles("sql/example.sql")

result$how_many_irises

result$n_longest_setosa_petal_lengths
```

Several aspects of SQL execution can also be controlled this way. See `vignette("sql_execution")` for complete details.

## Getting information
Several functions provide information about connections and the process of reading and preparing sql.
